Naive Bayes Classifier

### 2019-02-25 NB_2r2c.Rmd
update

---
title: "Naive Bayes Classifier"
date: "2/24/2019"
bibliography: library.bib
biblio-style: apalike
---

@@

```{}
#This is for convenience. It is not required if the number of training, test rows is known.
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

@@

Below is an **alternate version** of convert_function code block that KEEPs word frequencies while still using the same variable names and code structure. The processed matrices have cells with word frequencies instead of categorical yes/no values. This block only works for binary classification (ingroup/outgroup). It breaks when applied to data with multiple categories.
```{}
convert_counts2 <- function(x) {
  y <- x
}
```

Applies whichever "convert_counts2" function is currently active.
```{r apply_conversion}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

#Train Model, Predict, Evaluate for Subject
This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
```{r bayes_training}
# Train the classifier on training data, then test on test dataset, and full dataset.
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

@@

The faster method is to calculate a confusion matrix, which displays BOTH the truth table, and the final overall accuracy. A confusion matrix can be queried for specific values.
```{r reportOut}
#"line" is just spacer text for improving readability.
line<-("**************************************************************")

# Prepare the confusion matrix. 
conf.mat1 <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
conf.mat3 <- confusionMatrix(comments_test_pred_structure, comments_test_labels_structure)
conf.mat4 <- confusionMatrix(comments_all_pred_structure, comments_all_labels_structure)

#These commands show how to print out individual pieces of the confusion matrix. They show both the overall and by-group accuracy values for Subject x Test Subset.
conf.mat1
print (line)
print ("Test, Subject")

# This block outputs the by-group accuracy values in a single column for Subject x Test Subset. 
conf.mat1$byClass
print (line)
print ("Test, Subject")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x Test Subset.
conf.mat1$overall
print (line)
print ("Test, Subject")

# This block outputs JUST the overall accuracy value for Subject x Test Subset. 
conf.mat1$overall['Accuracy']
print ("Test, Subject")

# This block outputs both the overall and by-group accuracy values for Subject x All Data.
conf.mat2
print (line)
print ("All, Subject")

# This block outputs by-group accuracy values in a single column for Subject x All Data. 
conf.mat2$byClass
print (line)
print ("All, Subject")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x All Data.
conf.mat2$overall
print (line)
print ("All, Subject")

# This block outputs JUST the overall accuracy value for Subject x All Data. 
conf.mat2$overall['Accuracy']
print ("All, Subject")

# This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
conf.mat3
print (line)
print ("Test, Structure")

# This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
conf.mat3$byClass
print (line)
print ("Test, Structure")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
conf.mat3$overall
print (line)
print ("Test, Structure")

@@

# This block outputs both the overall and by-group accuracy values for Structure x All Data.
conf.mat4
print (line)
print ("All, Structure")

# This block outputs by-group accuracy values in a single column for Structure x All Data. 
conf.mat4$byClass
print (line)
print ("All, Structure")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
conf.mat4$overall
print (line)
print ("All, Structure")

@@

#Output for Review, Analysis
The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator. Results are exported to a CSV file. By default the CSV file should be saved to the current working directory.

@@

This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
```{}
tidy_prediction_table

@@

##Graph Outputs
Comparing outcome from Hand-coding vs. Bayes

***
Need to combine two datasets into one. 
***

```{r}
ggplot(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`,group=1))+
  geom_line()
ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
  geom_line()
ggplot(data=proptallysum3, aes(x=Hand.Coded.Structure, y=total, group=1))+
  geom_line()
ggplot(data=proptallysum4, aes(x=Bayes.Predicted.Structure, y=total, group=1))+
  geom_line()
```

>>>>

---
title: "Naive Bayes Classifier"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

@@

```{}
#This block is for convenience. It is not required if the number of training, test rows is known.
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

@@

The block below is an **alternate version** of the ```convert_function``` code block that KEEPs word frequencies while still using the same variable names and code structure. The processed matrices have cells with word frequencies instead of categorical yes/no values. This block **only** works when performing binary classification (ingroup/outgroup). The classifier fails when frequency values are used with data in multiple categories.
```{}
convert_counts2 <- function(x) {
  y <- x
}
```

This block applies whichever "convert_counts2" function is currently active.
```{r apply_conversion}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

#Train Model, Predict & Evaluate for Subject
This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
```{r bayes_training}
# Train the classifier on training data, then test on test dataset, and full dataset.
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

@@

The faster method is to calculate a confusion matrix, which displays BOTH the truth table, and the final overall accuracy. A confusion matrix can be queried for specific values.
```{r reportOut}
# Prepare the confusion matrix. 
conf.mat1 <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
conf.mat3 <- confusionMatrix(comments_test_pred_structure, comments_test_labels_structure)
conf.mat4 <- confusionMatrix(comments_all_pred_structure, comments_all_labels_structure)

#These commands show how to print out individual pieces of the confusion matrix. They show both the overall and by-group accuracy values for Subject x Test Subset.
print ("Test, Subject")
conf.mat1

# This block outputs the by-group accuracy values in a single column for Subject x Test Subset. 
print ("Test, Subject")
conf.mat1$byClass

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x Test Subset.
print ("Test, Subject")
conf.mat1$overall

# This block outputs JUST the overall accuracy value for Subject x Test Subset. 
print ("Test, Subject")
conf.mat1$overall['Accuracy']

# This block outputs both the overall and by-group accuracy values for Subject x All Data.
print ("All, Subject")
conf.mat2

# This block outputs by-group accuracy values in a single column for Subject x All Data. 
print ("All, Subject")
conf.mat2$byClass

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x All Data.
print ("All, Subject")
conf.mat2$overall

# This block outputs JUST the overall accuracy value for Subject x All Data. 
print ("All, Subject")
conf.mat2$overall['Accuracy']

# This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
print ("Test, Structure")
conf.mat3

# This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
print ("Test, Structure")
conf.mat3$byClass

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
print ("Test, Structure")
conf.mat3$overall

@@

# This block outputs both the overall and by-group accuracy values for Structure x All Data.
print ("All, Structure")
conf.mat4

# This block outputs by-group accuracy values in a single column for Structure x All Data. 
print ("All, Structure")
conf.mat4$byClass

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
print ("All, Structure")
conf.mat4$overall

@@

#Output for Review, Analysis
The code block below creates a CSV file containing the **original** unmodified TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator. Results are exported to a CSV file. By default the CSV file should be saved to the current working directory.

@@

This block is an alternative method to get the same data tables out using ```tidy``` format. It seems much less direct, but may be over-written.
```{r tidy_prediction_table}

@@

##Graph Outputs
This code block generates two graphs comparing the sub-category frequencies obtained by hand-coding vs. predicted frequencies from the Naive Bayes classifier. There are separate comparisons for the Subject and Structure categories.

```{r}
blendSubject = ggplot()+
  geom_col(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum2, aes(`Bayes.Predicted.Subject`, `total`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.1))+
  xlab("Subject Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comparison of Coded (blue) vs Predicted (red) Subject")

blendStructure = ggplot()+
  geom_col(data=proptallysum3, aes(`Hand.Coded.Structure`, `total`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum4, aes(`Bayes.Predicted.Structure`, `total`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.1))+
  xlab("Structure Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comparison of Coded (green) vs Predicted (orange) Structure")

print(blendSubject)
print(blendStructure)
```
